---
title: "Project 1 - Household Energy Expenditures - Electricity"
author: "Sean Franco, Mei Fei Hung, Daniel Felberg"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
#setwd() Meng Fei's WD  
setwd("C:/Users/18045/Documents/R/Data_Intro_Class/Project1")# Sean's WD
#setwd("C:/Users/danif/OneDrive/Documents/GWU - Data Science (Spring 2023)/DATS 6101/Project/Project1.R") Daniel's WD
library(readr)
library(ggplot2)
#install.packages("survey","ggmap","maps","mapdata","formattable", "forcats", "RColorBrewer","gridExtra", "usmap")
library(survey)
library(dplyr)
library(ggmap)
library(maps)
library(mapdata)
library(usmap)
library(formattable)
library(forcats)
library(RColorBrewer)
library(ezids)
library(readxl)
library(gridExtra)
library(lattice)
library(corrplot)
library(reshape2)
library(xtable)
library(car)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

Introduction:
This study examines the impact of household energy expenditure reported during the year 2015 from data from the Energy Information Administration. The study attempts to conduct exploratory data analysis (EDA) related to yearly household energy consumption, the expenditures of energy related to climate region, census income levels, and appliance types. The dataset for this project is from the 2015 survey from the Energy Information Administration which is the 14th edition of this national household survey. The year 2015 was chosen because it was before the coronavirus, and that it was after the 2008 financial crisis. Additionally, this survey is randomly sampled, and the most complete version of the survey with all amendments. 

Methods:
	For the most part, the attributes were pulled from the existing EIA survey data, and examined for normality. If heteroskedasticity was present, then leverage point outliers were removed to manipulate the data to normalize it. This could be a large assumption, but our rationale was that many of these outliers were not representative of the majority of the data points for the country. For example, a large outlier for total yearly electricity expenditure is likely to be from a very wealthy individual, let’s say the household for Jeff Bezos and family who can afford high prices for electricity, rather than any given working class family household. Almost all our variables needed 


```{r}
#LoadData and Example Code for Assigning weights----- 
#this is example code from the EIA weights doc:

RECS2015 <- read.csv("recs2015_public_v4.csv", header=TRUE, sep=",")

```
```{r}

app_cost <- data.frame(RECS2015$DOLELSPH,
                       RECS2015$DOLELCOL,
                       RECS2015$DOLELWTH,
                       RECS2015$DOLELRFG,
                       RECS2015$DOLELFRZ,
                       RECS2015$DOLELCOK,
                       RECS2015$DOLELMICRO,
                       RECS2015$DOLELCW,
                       RECS2015$DOLELCDR,
                       RECS2015$DOLELDWH,
                       RECS2015$DOLELLGT,
                       RECS2015$DOLELTVREL,
                       RECS2015$DOLELCFAN)

colnames(app_cost) <- c(
  "Spaceheating", 
  "AC", 
  "Waterheating",
  "Refrigerator", 
  "Freezer",
  "Cooking", 
  "Microwave",
  "Clothewasher", 
  "Clothedryer",
  "Dishwasher",
  "Light", 
  "TV",
  "Fans")

app_cost <- outlierKD2(app_cost, (Spaceheating), rm =TRUE)
app_cost <- outlierKD2(app_cost, (AC), rm =TRUE) 
app_cost <- outlierKD2(app_cost, (Waterheating), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Light), rm =TRUE)
app_cost <- outlierKD2(app_cost, (TV), rm =TRUE)  
app_cost <- outlierKD2(app_cost, (Dishwasher), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Refrigerator), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Freezer), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Cooking), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Microwave), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Clothewasher), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Clothedryer), rm =TRUE)
app_cost <- outlierKD2(app_cost, (Fans), rm =TRUE)

app_cost <- app_cost[-ncol(app_cost)]

app_cost["series"] <- rep(1, nrow(app_cost))

test <- melt(app_cost,  id.vars = 'series', variable.name = 'index')

test %>%
  arrange(index)%>%
  mutate(index = factor(index, levels = c("series",
                                          "Waterheating",
                                          "AC",
                                          "Spaceheating", 
                                          "Light",
                                          "TV",
                                          "Refrigerator",
                                          "Clothedryer",
                                          "Fans",
                                          "Freezer",
                                          "Cooking", 
                                          "Microwave", 
                                          "Dishwasher",
                                          "Clothewasher"
                                          ))) %>%
  ggplot(
    aes(x = series,
        y = value,
        fill = index))+ 
  geom_bar(stat = "identity", 
           position = "dodge") +
  labs(title = "Annual Electricity Cost of Different Appliances", 
       x = "Appliances",
       y = "Annual Electricity Cost")+  
  theme(axis.text.x = element_blank())
```
See above is a plot of common appliances and their total yearly electricity usage. We can notice that waterheating, spaceheating and air conditioning, are among the top appliances that consume electricity. Examining central air costs for heating and cooling are among the top energy expenses for households. 

```{r}

RECS2015$DOLELSPH <- currency(RECS2015$DOLELSPH,
                              symbol = "$",
                              digits = 2L,
                              format = "f",
                              big.mark = ",",
                              sep = "")

RECS2015$DOLELCOL <- currency(RECS2015$DOLELCOL,
                              symbol = "$",
                              digits = 2L,
                              format = "f",
                              big.mark = ",",
                              sep = "")

RECS2015 <- RECS2015 %>% 
  mutate(CENACHP = as.factor(case_when(CENACHP == 1 ~ "Has a Heat Pump",
                                       CENACHP == 0 ~ "No Heat Pump",
                                       CENACHP == -2 ~ "NA")))

central_air_df <- data.frame(RECS2015$DOLELSPH, 
                             RECS2015$DOLELCOL, 
                             RECS2015$CENACHP)

colnames(central_air_df) <- c("Electricity Space Heating Costs", 
                              "Electricity AC Costs", 
                              "Heat Pump Status")

#dataframe prep for combo plot
i <- 1
newdf <- 0
for(i in 2:2){
  
  newdf["Electricity Costs in $"] <- rbind(as_tibble(central_air_df[,1]), as_tibble(central_air_df[,i]))
  
}
newdf <- as.data.frame(newdf)
newdf["Heat Pump Status"] <- rep(central_air_df$`Heat Pump Status`, 2) 
newdf["Group"] <- rbind(as_tibble(rep("Heating", nrow(central_air_df))), 
                        as_tibble(rep("Cooling", nrow(central_air_df)))) 

newdf <- newdf[-1]
colnames(newdf) <- c("Electricity Costs in $", "Heat Pump Status", "Group")

group.colors <- c("Heating" = "Orange", "Cooling" = "Light Blue")

ggplot(newdf,
       aes(x = `Heat Pump Status`,
           y = `Electricity Costs in $`,
           fill = `Group`))+
  geom_col(stat = "identity", position = "dodge")+
  labs(title = "Yearly Central Air Costs (in $)")+
  scale_fill_manual(values=group.colors)
```

This plot shows heating and cooling costs for households with heat pumps and without heatpumps. Households with heatpumps roughly save half of their heating costs than households without heat pumps. A T-test (see below) indicated that the difference between households with heatpumps is significant for total, yearly heating costs.

Cooling costs for households with heat pumps save almost $500 a year (in total cooling costs) compared to households without heat pumps. A T-test (see below) indicated that the difference between households with heatpumps is significant for total, yearly AC costs.

```{r}
#T-Test for heatpumps v. spaceheater----

ttest_central_air_df <- central_air_df %>%
  mutate(`alt` = central_air_df$`Heat Pump Status` == "Has a Heat Pump",
         'null' = central_air_df$`Heat Pump Status` == "No Heat Pump")

no_hp <- ttest_central_air_df %>%
  filter(null == TRUE)
    
has_hp <- ttest_central_air_df %>%
  filter(alt == TRUE)

t.test(x = has_hp$`Electricity Space Heating Costs`,
       conf.level = 0.95,
       mu = mean(no_hp$`Electricity Space Heating Costs`))

#T-Test for heatpumps v. AC-----

no_hp <- ttest_central_air_df %>%
  filter(null == TRUE)

has_hp <- ttest_central_air_df %>%
  filter(alt == TRUE)

t.test(x = has_hp$`Electricity AC Costs`,
       conf.level = 0.95,
       mu = mean(no_hp$`Electricity AC Costs`))

```

```{r}
#----------------------Income and Energy Expenditure info----------------------- ------
#Annual gross household income for the last year (“MONEYPY” variable),

RECS2015 <- RECS2015 %>%
  mutate(MONEYPY = as.factor(case_when(MONEYPY == 1 ~ "Less than $20,000",
                                       MONEYPY == 2 ~ "$20,000 - $39,999",
                                       MONEYPY == 3 ~ "$40,000 - $59,999",
                                       MONEYPY == 4 ~ "$60,000 to $79,999",
                                       MONEYPY == 5 ~ "$80,000 to $99,999",
                                       MONEYPY == 6 ~ "$100,000 to $119,999",
                                       MONEYPY == 7 ~ "$120,000 to $139,999",
                                       MONEYPY == 8 ~ "$140,000 or more")))

central_air_df["Income"] <- data.frame(RECS2015$MONEYPY)

has_hp <- central_air_df %>%
  filter(`Heat Pump Status` == "Has a Heat Pump")

no_hp <- central_air_df %>%
  filter(`Heat Pump Status` == "No Heat Pump")

central_air_df %>%
  arrange(`Heat Pump Status`) %>%
  mutate(Income = factor(Income, levels = c("Less than $20,000",
                                            "$20,000 - $39,999",
                                            "$40,000 - $59,999",
                                            "$60,000 to $79,999",
                                            "$80,000 to $99,999",
                                            "$100,000 to $119,999",
                                            "$120,000 to $139,999",
                                            "$140,000 or more")))%>%
  ggplot(aes(x = Income,
             y = `Heat Pump Status`,
             fill = `Heat Pump Status`))+
  geom_bar(stat = "identity")+
  labs(title = "Income Bracket And Heat Pump Status",
       ylab = "")+
  theme(axis.text.x = element_text(angle = 45, size = 9, margin = margin(r=0)),
        axis.text.y=element_blank()) #needs to fix income brackets to be ascending

```

See above is a bar plot of Heat pump status as binary counts based on household income brakets. Based on ocular scrutiny, higher income brackets report larger ratios of having heatpumps. The $20k - $39k income bracket have the highest counts of not having heatpumps, and having heat pumps. The $40k - $59k; $140k to more income brackets have high ratios of having heat pumps.

#put anova test here

```{r}

#BOXPLOT of income and yearly energy costs----
central_air_df["TotElectricity"] <- RECS2015$DOLLAREL

central_air_df <- outlierKD2(central_air_df, TotElectricity, rm= TRUE)

#PLOT income by total electricity usage-----
central_air_df %>%
  arrange(`Heat Pump Status`) %>%
  mutate(Income = factor(Income, levels = c("Less than $20,000",
                                            "$20,000 - $39,999",
                                            "$40,000 - $59,999",
                                            "$60,000 to $79,999",
                                            "$80,000 to $99,999",
                                            "$100,000 to $119,999",
                                            "$120,000 to $139,999",
                                            "$140,000 or more")))%>%
  ggplot(aes(x = Income,
             y = TotElectricity, 
             color = Income)) + 
  geom_boxplot(stat = "boxplot") +
  labs(title = "Yearly Electricity Cost by Income Level ", 
       x = "Income level",
       y = "Yearly Electricity Cost (in $)")+  
  theme(axis.text.x = element_text(angle = 45, size = 9, margin = margin(r=0)), 
        axis.text.y=element_text())
```

See above are boxplots of income brackets and total yearly electricity costs. On averge, lower incomes tend to spend less on electricity costs, compared to higher incomes. This is also supported in literature about energy poverty and cultural practices to save energy. 

#put Anova test here

```{r}
#Spatial Prep------

#Preparing Census Regions 

RECS2015 <- RECS2015 %>%
  mutate(REGIONC = as.factor(case_when(REGIONC == 1 ~ "Northeast",
                                       REGIONC == 2 ~ "Midwest",
                                       REGIONC == 3 ~ "South",
                                       REGIONC == 4 ~ "West")))
RECS2015$REGIONC
plot(RECS2015$REGIONC)

#Preparing Census Divisions

RECS2015 <- RECS2015 %>%
  mutate(DIVISION = as.factor(case_when(DIVISION == 1 ~ "New England",
                                       DIVISION == 2 ~ "Middle Atlantic",
                                       DIVISION == 3 ~ "East North Central",
                                       DIVISION == 4 ~ "West North Central",
                                       DIVISION == 5 ~ "South Atlantic",
                                       DIVISION == 6 ~ "East South Central",
                                       DIVISION == 7 ~ "West South Central",
                                       DIVISION == 8 ~ "Mountain North",
                                       DIVISION == 9 ~ "Mountain South",
                                       DIVISION == 10 ~ "Pacific",)))

#Electricity cost differences between urban and rural areas (“UATYP10” variable),

#rename levels of area 
RECS2015 <- RECS2015 %>% 
  mutate(UATYP10 = as.factor(case_when(UATYP10 == "U" ~ "Urban Area",
                                       UATYP10 == "R" ~ "Rural",
                                       UATYP10 == "C" ~ "Urban Cluster")))

#dataframe of area and total energy
Tot_Energy_area_df <- data.frame(RECS2015$UATYP10, RECS2015$DOLLAREL, RECS2015$DIVISION, RECS2015$CLIMATE_REGION_PUB)
colnames(Tot_Energy_area_df) <- c("Urban Density", "Yearly Electricity Costs", "Division", "Climate")

#HISTOGRAM and Q-Q Plot before outliers are removed----
ggplot(data=Tot_Energy_area_df, aes(x = `Yearly Electricity Costs`)) + 
  geom_histogram(breaks=seq(19, 8122, by = 100), 
                 col="black", 
                 fill="dark green", 
                 alpha = .7) + # opacity
  labs(x="Electricity Cost", y="Frequency") +
  labs(title="Histogram of Total Electricity Cost, Using `ggplot`")

qqnorm(Tot_Energy_area_df$`Yearly Electricity Costs`, main = "Q-Q Plot of Total Electricity Cost")
qqline(Tot_Energy_area_df$`Yearly Electricity Costs`)

### Removing outliers:
Tot_Energy_area_df <- outlierKD2(Tot_Energy_area_df, `Yearly Electricity Costs`, rm= TRUE)

#HISTOGRAM and Q-Q Plot again after removing outliers----

ggplot(data=Tot_Energy_area_df, aes(x = `Yearly Electricity Costs`)) + 
  geom_histogram(breaks=seq(19, 8122, by = 100), 
                 col="black", 
                 fill="dark green", 
                 alpha = .7) + # opacity
  labs(x="Electricity Cost", y="Frequency") +
  labs(title="Histogram of Total Electricity Cost, Using `ggplot`")

qqnorm(Tot_Energy_area_df$`Yearly Electricity Costs`, main = "Q-Q Plot of Total Electricity Cost")
qqline(Tot_Energy_area_df$`Yearly Electricity Costs`)

```

```{r}

#PLOT Yearly expenditure differences by Division and Density-----
Tot_Energy_area_df %>%
  arrange(`Yearly Electricity Costs`)%>%
  mutate(Division = factor(Division, levels = c("New England",
                                                "Middle Atlantic",
                                                "East North Central",
                                                "West North Central",
                                                "South Atlantic",
                                                "East South Central",
                                                "West South Central",
                                                "Mountain North",
                                                "Mountain South",
                                                "Pacific")))%>%
  ggplot(
    aes(x = `Urban Density`,
        y = `Yearly Electricity Costs`,
        fill = fct_reorder(`Division`, `Yearly Electricity Costs`)))+
  geom_boxplot(stat = "boxplot", 
               position = "dodge")+
  labs(title = "Urban Density and Electricity Costs",
       fill= "Division")+
  theme(axis.text.x = element_text(size = 9, margin = margin(r=0)), 
        axis.text.y=element_text())+ 
  scale_color_brewer(palette = "Pastel2")

```

See above is a boxplot of census divisions and total yearly electricity costs. We can observe that households in rural areas spend more on electricity than Urban or Urban Clusters. We can also observe that urban areas genereally have high populations which may explain the more distant outliers. 